### 2014-08-15 drafts/ApproximatePCFG/ApproximatePCFG.tex
Initial draft commits.

\documentclass{article}

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
\usepackage{amsmath}

\usepackage{algorithm}
\usepackage{algpseudocode}

\author{Eyal Dechter}

\begin{document}
The Inside-Outside algorithm is a popular algorithm for computing the
MAP marginals of PCFG rules given a sentence or collection of
sentences. The algorithm's computational complexity is $O(n^3 K)$
where $n$ is the length of the sentence and $K$ is the the number of
rules in the PCFG. The cubic algorithm makes inference tractable for
relatively long sentences -- containing perhaps a few dozen terminals
-- but makes it impossible for this algorithm to be used in an online
fashion, or to be used on grammars that span multiple levels of
linguistic abstration (for example, imagine trying to learn
morphological and syntactic grammar at the same time with the
Inside-Outside algorithm).

Our goal here is to develop an algorithm that approximates the MAP
marginals of a PCFG given data but that can run much faster; further,
we would like to be able to put bounds on the approximation. The core
intuition is that if we know that the probability of some symbol
dominating a span cannot be more than some very small amount, we
shouldn't have to calculate that amount accurately; instead, we should
be able to substitute this small upperbound in our calculations,
absolving us of exploring parts of the space of parses that are very
unlikely to be relevant.

In what follows we will assume that we have a PCFG $G$ in Chomsky
normal form with start symbol $S$, non-terminals $\mathcal{A}$,
terminals $\mathcal{T}$, and rules $\mathcal{R}$, where we associate
with each rule $r \in \mathcal{R}$ a potential $\psi(r)$. We also
assume that we are given a sentence $x = x_1, \dots, x_N$.

For every triple $(A, i, j)$ where $A$ is a nonterminal and $1 \leq i
\leq j \leq N$, the Inside-Outside algorithm computes an alpha and
beta value. The alpha and beta values are defined according to the
following recurrence:

\begin{align}
\alpha(A, i, i) &= \psi(A \rightarrow x_i)\\
\alpha(A, i, j) &= \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)\\
\beta(A, 1, N) &= 1 \\
\beta(A, i, j) &= \sum_{k=1}^{i-1} \sum_{B \rightarrow C A \in \mathcal{R}}
  \beta(B, k, j) \psi(B \rightarrow C A) \alpha(C, k, i-1)\\ 
  &+  \sum_{k=j+1}^{N} \sum_{B \rightarrow A C \in \mathcal{R} : A \neq C}
  \beta(B, i, k) \psi(B \rightarrow A C) \alpha(C, j+1, k) 
\end{align}

Suppose we have an approximate $\alpha'$ of $\alpha$ with the
property that for some small $\epsilon > 0$ and for all $A, i, j$:

\begin{align}
\alpha'(A, i, i) &= \alpha(A, i, i)\\ 
\alpha(A, i, j) &\leq \alpha'(A, i, j) \leq (1+\epsilon)  \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
  \psi(A \rightarrow B C ) \alpha'(B, i, k) \alpha'(C, k+1, j).\\
\end{align}

By how much can $\alpha'$ over estimate $\alpha$? Let $E(A, i,
j)=\frac{\alpha'(A, i, j)}{\alpha(A, i, j)}$ be the measure of this
over-estimate. We have that

\begin{align}
E_{\alpha}(A, i, i) &= 1\\
E_{\alpha}(A, i,j) &= \frac{\alpha'(A, i, j)}{\alpha(A, i, j)}\\
          &\leq (1+\epsilon) \frac{ \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
  \psi(A \rightarrow B C ) \alpha'(B, i, k) \alpha'(C, k+1, j)}
           {\sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)}\\
          &=  (1+\epsilon) \frac{ \sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}} 
  \psi(A \rightarrow B C ) E_{\alpha}(B, i, k) \alpha(B, i, k) E_{\alpha}(C, k+1, j) \alpha(C, k+1, j)}
           {\sum_{k=i}^{j-1} \sum_{A \rightarrow B C \in \mathcal{R}}
  \psi(A \rightarrow B C ) \alpha(B, i, k) \alpha(C, k+1, j)}\\
         &= (1+\epsilon) \max_{k\in \{i, \dots, j-1\}, A \rightarrow B C \in \mathcal{R} }
            E_{\alpha}(B, i, k) E_{\alpha}(C, k+1, j).
\end{align}

Let $\hat{E}_{\alpha}(d) = \max_{A\in\mathcal{A}, i \in \{1, \dots, N-d \}}
E_{\alpha}(A, i, i+d)$ be the maximum upper bound above over all spans of
length $d$. Then we have that

\begin{align}
\hat{E}_{\alpha}(1) &= 1\\
\hat{E}_{\alpha}(d) &= (1+\epsilon) \max_{s \in \{1, \dots, d-1\}} \hat{E}_{\alpha}(s) \hat{E}_{\alpha}(d-s).\\
\end{align}

Solving this recurrence, gives $\hat{E}_{\alpha}(d) = (1 + \epsilon)^{d-1}$. To
re-iterate, what this means is that if we have such an $\alpha'$ then
for all $A, i, j$, $\alpha'(A, i, j) \leq (1 + \epsilon)^{j - i}
\alpha(A, i, j)$.

We can now develop an analog $\beta'$ of $\beta$ which has the same
dependency on $\alpha'$ that $\beta$ has on $\alpha$. Let
$\hat{E}_{\beta}(d)$ be an upper bound on $\frac{\beta'(A, i,
  i+d)}{\beta(A, i, i+d)}$ for all $A$ and $i$. Then a similar
analysis as above shows that one such $\hat{E}_\beta$ is

\begin{align}
\hat{E}_{\beta}(N) &= 1\\
\hat{E}_{\beta}(N-k) &= (1+\epsilon)^{N-k-1} \max_{s \in \{N-k+1, \dots, N\}} \hat{E}_{\beta}(s).
\end{align}

Solving this recurrence yields $\hat{E}_{\beta}(d) = (1+\epsilon)^{\frac{(N-d)(N+d-1)}{2}}$.

In the Inside-Outside algorithm, the merit $\mu(A, i, j) = \alpha(A,
i, j) \beta(A, i, j)$ is the unnormalized probability that nonterminal
$A$ yields span $(i, j)$. Define $\mu'(A, i, j) = \alpha'(A, i, j)
\beta(A, i, j)$. By the results above, we have that the ratio of the
actual merit to our approximation is upper bounded by
$(1+\epsilon)^{\frac{(N+d-2)(N-d+1)}{2}}$, where, as before, $d$ is
the length of the span $(i, j)$.

\section{Approximate Inside}
Here we describe the algorithm ApproxInside which computes the
$\alpha'$ values.

\begin{algorithm}
\caption{The Approximate Inside Algorithm \label{alg:approxInside}}
\begin{algorithmic}[1]
\Procedure {ApproxInside}{$s$, $G=(S, bRules, uRules)$, $U$, $\epsilon$}
  \State \Comment $s$ sentences of length $n$
  \State \Comment $G$ grammar 
  \State \Comment $S$ start symbol
  \State \Comment $U$ upper bound, function from $(A, i, j) \mapsto \mathbf{R}$
  \State $\alpha \leftarrow \Call{InitAlpha}{G, 1, n}$ \Comment Initialize table of $\alpha$ 
  \State \Return \Call{$GO_{OR}$}{$S$, $1$, $n$}
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{The Approximate Inside Algorithm (cont)}
\begin{algorithmic}[2]
\Procedure {$GO_{OR}$}{$A$, $i$, $j$}
  \State \Comment $A$ nonterminal symbol
  \State $children \leftarrow \Call{GetChildren}{A, i, k}$
  \State $lb \leftarrow 0 $ \Comment Initialize lower bound on OR node to 0
  \For{$node \in children$}{}{}
    \If{$node$ is unary of form $A \rightarrow t, i, i$ and $t = s_i$}
      \State $v \leftarrow \psi(A \rightarrow t)$
    \ElsIf{$node$ is binary}
      \State $v \leftarrow $\;\Call{$GO_{AND}$}{$node$, $lb$}
    \EndIf
    \State $lb \leftarrow lb + v$ \Comment Update lower bound
  \EndFor
  \State $\alpha[A, i, j] \leftarrow lb$ \Comment Insert $\alpha$ value into table.
\EndProcedure
\Statex
\Procedure {$GO_{AND}$}{$A \rightarrow B\, C$, $i$, $k$,$j$, $lb$}
  \State $w \leftarrow \psi(A \rightarrow B\, C)$ 
  \State $v_{\ell} \leftarrow \alpha[B, i, k]$ 
  \State $v_{r} \leftarrow \alpha[C, k+1, j]$
  \State $u_{\ell} \leftarrow U[B, i, k]$ 
  \State $u_{r} \leftarrow U[C, k+1, j]$
  \If{neither $v_{\ell}$ nor $v_r$ is $NIL$} \Comment If both values are already computed, just use them
    \State \Return $w * v_{\ell} * v_r$
  \ElsIf{ $v_r \neq NIL$}  \Comment If the right value is already computed, compute left value
    \If {$u_\ell \leq \frac{\epsilon * lb}{w * v_r}$}
      \State $v_\ell \leftarrow u_\ell$
    \Else
      \State $v_\ell \leftarrow $\;\Call{$GO_{OR}$}{$B$, $i$, $k$}
    \EndIf
  \ElsIf{ $v_\ell \neq NIL$}  \Comment If the left value is already computed, compute right value
    \If {$u_r \leq \frac{\epsilon * lb}{w * v_\ell}$}
      \State $v_r \leftarrow u_r$
    \Else
      \State $v_r \leftarrow $\;\Call{$GO_{OR}$}{$C$, $k+1$, $j$}
    \EndIf
  \Else \Comment If neither values yet computed, compute both
    \If {$u_\ell u_r \leq \frac{\epsilon * lb}{w}$}
      \State $v_r \leftarrow u_r$
      \State $v_\ell \leftarrow u_\ell$
    \Else
      \State $v_\ell \leftarrow $\;\Call{$GO_{OR}$}{$B$, $i$, $k$}
      \State $v_r \leftarrow $\;\Call{$GO_{OR}$}{$C$, $k+1$, $j$}
    \EndIf
  \EndIf
  \State \Return $w * v_{\ell} * v_r$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{The Approximate Outside Algorithm.\label{alg:approxOutside}}
\begin{algorithmic}[2]
\Procedure {$ApproxOutside$}{$s$, $G$, $S$}
  \State $n \leftarrow length(s)$
  \State $\beta[S, 1, n] \leftarrow 1$ \Comment initialize $\beta$ recursion, unless $n==1$ in which case we are done
  \State $open \leftarrow [(S, 1, n)]$ \Comment initialize list that will hold nodes to be processed
  \While {$open \neq NIL$}
  \State $children \leftarrow []$ \Comment initialize list that holds next level of nodes to be processed
   \For {$(A, i, j) \in open$ \Comment $i \neq j$}
    \For {$A \leftarrow B\, C \in rules(G)$} \Comment iterate over binary rules headed by $A$
      \For {$k \leftarrow i, \dots, j - 1$}
        \If {$\alpha[B, i, k] \neq 0$}
          \State if $\beta[C, k+1, j] = NIL$, set to $0$
          \State $\beta[C, k+1, j] \leftarrow \beta[C, k+1, j] + \psi(A \rightarrow B C) \alpha[B, i, k] \beta[A, i, j]$
          \State add $(C, k+1, j)$ to $children$
        \EndIf

        \If {$\alpha[C, k+1, j] \neq 0$}
          \State if $\beta[B, i, k] = NIL$, set to $0$
          \State $\beta[B, i, k] \leftarrow \beta[B, i, k] + \psi(A \rightarrow B C) \alpha[C, k+1, j] \beta[A, i, j]$
          \State add $(B, i, k)$ to $children$
        \EndIf 
      \EndFor 
   \EndFor
   \EndFor
   \State $open \leftarrow children$
  \EndWhile    
\EndProcedure
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{The Approximate Outside Algorithm.\label{alg:approxOutside}}
\begin{algorithmic}[2]
\Procedure {$ApproxOutside$}{$s$, $G$, $S$}
  \State $OPEN \leftarrow []$ \Comment initialize OPEN FIFO queue
  \State $n \leftarrow length(s)$
  \State $\beta(S, 1, n) \leftarrow 1$
  \State \Call{AddChildrenToOpen}{$S, 1, n, \beta(S, 1, n)$}
  \While {$OPEN$ is not empty}
    \State $node \leftarrow pop(OPEN)$ where $rule_{node} = (A \rightarrow B C, i, k, j) = fst(node)$ and $pa_\beta= snd(node)$
    \State $a_\ell \leftarrow \alpha[B, i, k]$
    \State $a_r \leftarrow \alpha[C, k+1, j]$
    \If {$a_r \neq NIL$}
      \If {$\beta[B, i, k] = NIL$}
         $\beta[B, i, k] \leftarrow 0$
      \EndIf
      \State $\beta[B, i, k] \leftarrow \beta[B, i, k] + pa_\beta a_\ell \psi(A \rightarrow B C)$ \Comment compute beta value and add to current beta value 
      \State \Call{AddChildrenToOpen}{$B, i, k, \beta(B, i, k)$}
    \EndIf
    \If {$a_l \neq NIL$}
      \If {$\beta[C, k+1, j] = NIL$}
         $\beta[C, k+1, j] \leftarrow 0$
      \EndIf
      \State $\beta[C, k+1, j] \leftarrow \beta[C, k+1, j] + pa_\beta a_r \psi(A \rightarrow B C)$ \Comment compute beta value and add to current beta value 
      \State \Call{AddChildrenToOpen}{$C, k+1, j, \beta(C, k+1, j)$}
    \EndIf
  \EndWhile  
\EndProcedure

\Statex
\Procedure {AddChildrenToOPEN}{A, i, j, betaVal}
  \State $\{c_i\}\leftarrow \Call{GetChildren}{A, i, j}$ \Comment Get AND-node children
  \State insert pairs $\{(c_i, betaVal\}_i)$ into OPEN
\EndProcedure
\end{algorithmic}
\end{algorithm}

\noindent \textbf{Claim: Algorithm~\ref{alg:approxInside} correctly computes the $\beta'$ values.}
\noindent TODO

\noindent \textbf{Claim: Algorithm~\ref{alg:approxOutside} correctly computes the $\beta'$ values.}
\noindent TODO

\section{A simple upper bound}
We will present a simple upper bound on a slightly restricted set of
PCFGs which is based on the following intuition: if I am a nonterminal
$A$ and I know that I need to yield a string $x_1 \dots x_n$, then
this is going to be at least as hard as starting at $A$ and generating
any one of the elements in $x_1 \dost x_n$. That is, the highest
probability single word in my yield will have a higher probability of
being produced, then any longer span.

Suppose our PCFG has potentials less than or equal to one. This is the
case in the commons situation in which the PCFG is normalized, or if,
as in variational approximations to the Hierarchical Diirichlet PCFG,
our PCFG is unnormalized with values summing to less than one. Now we
define a function $g(A, w, m)$ for nonterminal $A$, terminal $w$ and
integer $w>0$, which we claim is greater the weight our PCFG would
assign to any parse tree rooted at $A$ and yielding any span of length
$m$ containing $w$. Specifically, consider any such parse tree, and
set all weights not on the path between $A$ and $w$ to 1. The total
weight assigned to the parse tree is the product of the weights on its
edges, so the resulting parse tree will have a strictly greater score.

\begin{algorithm}
\caption{An O(n^2}
\begin{algorithmic}[2]
\Procedure {$GO_{OR}$}{$A$, $i$, $j$}

\EndProcedure
\end{algorithmic}
\end{algorithm}

\end{document}