### 2014-11-07 drafts/nips_workshop/nips_workshop.tex

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{anyfontsize}
\usepackage{subcaption}
\usepackage{titling}
\usepackage{array}
\usepackage{authblk}

\title{\vspace{-1.5cm} Unsupervised learning of probabilistic programs with latent predicate networks}

\author{Eyal Dechter\thanks{{\tt \{edechter, rule, jbt\} @mit.edu}}}
\author{Joshua Rule}
\author{Joshua B. Tenenbaum}
\affil{{\normalsize Department of Brain and Cognitive Sciences, MIT}}
\date{}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

\vspace{-1cm}
\begin{abstract}

  Probabilistic programs successfully capture many aspects of the
  complex symbolic and statistical knowledge brought to bear in human
  cognition. Children aren't born with this knowledge, however, but
  must instead learn it, often in an unsupervised fashion. Compelling
  probabilistic programming models of human cognition must therefore
  also be learned. We present early work on Latent Predicate Networks
  (LPN), a framework for unsupervised learning in a restricted space
  of context-sensitive probabilistic programs. Specifically, we use
  Online Variational Bayesian Expectation Maximization in
  PRISM, a probabilistic programming system, to perform inference over
  the set of Stochastic Logic Programs (SLP) parseable in polynomial
  time. Using a small fragment of English sentences about number, we
  show that LPNs can recover the relevant conceptual structure
  necessary to generate novel sentences in this domain with high
  probability.

\end{abstract}

\section{Introduction}

%% Introduction idea
%% *** Probabilistic programs have been great models of human cognition
%% **** one-shot learning
%% **** linguistic pragmatics
%% **** some other example
%% *** one thing they lack is a good way to automatically induce full programs from data
%% *** yet, children seem to do this all the time
%% *** So, this motivates our approach
%% *** But, we have to overcome some technical challenges
%% **** many prob.prog. languages are too expressive to be tractable
%% **** many models of concept learning use too restricted a set of primitives

Probabilistic programs combine symbolic and statistical knowledge in way that is
natural for humans to express. In fact, probabilistic programs have
been proposed as models of human cognition in a variety of domains
[CITATION]. As a suitable representation for learning, however, the flexibility
of probabilistic programs poses a challenge. At one extreme, we can
approach learning as fixing the symbolic structure of a probabilistic
program and fitting some of its parameters; at the other extreme, the
structure is unspecified and learning is performed by searching over
valid syntactic forms. The former sacrifices the symbolic flexibility
that probabilistic programs have to offer, and the latter is
computationally infeasible. It is natural to suppose, then, that human
learning charts a middle course between these extremes, restricting
the space of probabilistic programs -- perhaps in a domain-specific
way -- to accomodate both flexibility and tractability.

Here, we present statistical learning (using the PRISM probabilistic
logic programmming language) of a restricted type of probabilistic
program, dubbed a \emph{latent predicate network} (LPN), for learning
problems at the intersection of concept learning and language
acquisition. People learn many abstract concepts primarily through
language, and understanding language depends on understanding the
underlying concepts. Number concepts are a good example: children do
not learn about the meaning of ``seventy five'' by seeing examples of
seventy five things; they do not know that ``seventy five'' is more
than ``twenty five'' because of their perceptual experiences of these
quantities. Rather, children learn the meaning of ``seventy five'' (or
``a billion and five'') by noticing how number words are used in
language, in counting sequences, in arithmetic exercises, {\it
  etc}. Other good examples of such abstract concepts are kinship and
social relations ({\it e.g.} ``my father in law's grandmother''),
temporal relations (``the day after last Thanksgiving.''), and spatial
relations (``above and just to the left of'').

We use LPNs in an attempt to address the issues of tractibility and
symbolic flexibility. We address the issue of tractability by
extending Range Concatenation Grammars -- a context-sensitive grammar
formalism originally developed in linguistic -- with a probabilistic
interpretation, resulting in probabilistic programs in which
inference, and thus learning, is relatively tractable. We address the
issue of structural flexibility by building in latent predicates,
which, through learning, assume the role of a domains underlying concepts (in the
number domain, these might correspond to the concepts of
successorship, order of magnitude, magnitude comparison, exact
vs. approximate, {\it etc}.).

% Such concepts share many of the properties of language syntax: they
% are unbounded in number, they derive their meanings via composition,
% and, although people only ever say, hear, or read about a small number
% of them, they are able to reason correctly about any them. There seems
% to be a grammar to these concepts, and grasping this grammar is
% critical to understanding their meanings and how to use them. This
% motivates our approach here, which is to apply the tools of
% probabilistic grammars more familiar from studies of syntax to the
% problem of concept aquisition.

% {\bf include somewhere:} A general probabilistic logic
% programming system based on PROLOG, PRISM provides built-in predicates
% for probabilistic execution and Bayesian inference over logic programs
% with stochastic choices.

% Doing so requires overcoming some technical barriers.

% First, many formal systems of probabilistic programming are so
% expressive as to make learning generally intractable. To address this,
% we use Range Concatenation Grammar, a context-sensitive grammar
% formalism originally developed in linguistics and extend this to a
% probabilistic model of programs.

% Second, the categories of syntax -- the nonterminals of the grammar --
% are often assumed to be known to children innately and given to
% automated learners by human experts. The categories that underlie
% conceptual knowledge, on the other hand, are far more numerous, vary
% from domain to domain, and are unlikely to be known to the
% learner. This motivates our use of latent predicates that, through
% learning, assume the role of a domain's underlying concepts (in the
% number domain, these might correspond to the concepts of
% successorship, order of magnitude, magnitude comparison, exact
% vs. approximate, {\it etc}).

% Finally, inducing probabilistic context-sensitive grammars with latent
% predicates threatens to be intractable: our goal is to find a middle
% ground between expressivity and tractability. Using
% PRISM~\cite{DBLP:journals/jair/SatoK01} -- a probabilistic logic
% programming system that naturally implements efficient dynamic
% programming algorithms for our models -- we are able to explore which
% domains and which grammar architectures are a good fit for this
% grammar-based approach.

The rest of this paper describes early work on this approach. First,
we present Latent Predicate Networks (LPNs). Then, we describe our
approach to inference and its implementation in PRISM. Finally, we
present preliminary experimental results in the domain of number
concept learning.

\section{Latent Predicate Networks}

\begin{figure}[t]
  \begin{subfigure}[b]{0.5\linewidth}
    \includegraphics[width=\linewidth]{lpn/lpn.pdf}
    \caption{The architecture and rules of a schematic LPN.}
    \label{fig:architecture}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.5\linewidth}
    \includegraphics[width=\linewidth]{parseTree/parse.pdf}
    \caption{A possible parse of the sentence ``after twenty five comes twenty six" using a 5-predicate LPN.}
    \label{fig:parseexample}
  \end{subfigure}
\end{figure}

An LPN is a hierarchical Bayesian model of strings
extending the Hierarchical Dirichlet PCFG model (HD-PCFG) to
Probabilistic Range Concatenation Grammars (PRCGs).

\subsection{Probabilistic Range Concatenation Grammars}
Range Concatenation Grammars (RCGs) are a class of string grammars
that represent all and only those languages that can be parsed in time
polynomial in the length of the target
string~\cite{boullier2005range}. An RCG $G=(N, T, V, P, S)$ is a
5-tuple where $N$ is a finite set of predicate symbols, $T$ is a set
of terminal symbols, $V$ is a set of variable symbols, P is a finite
set of $M \geq 0$ clauses of the form $\psi_0 \rightarrow \psi_1 \dots
\psi_M$, and $S \in N$ is the \emph{axiom}. Each $\psi_m$ is a term of
the form $A(\alpha_1, \dots, \alpha_{\mathcal{A}(A)}$, where $A \in
N$, $\mathcal{A}(A)$ is the arity of $A$, and each $\alpha_i \in (T
\cup V)^*$ is an argument of $\psi_m$. We call the left hand side term
of any clause the \emph{head} of that clause and its predicate symbol
is the \emph{head predicate}.

A string $x$ is in the language defined by an RCG if one can
\emph{derive} $S(x)$. A derivation is a sequence of rewrite steps in
which substrings of the left hand side argument string are bound to
the variables of the head of some clause, thus determining the
arguments in the clause body. If a clause has no body terms, then its
head is derived; otherwise, its head is derived if its body clauses
are derived.\footnote{This description of the language of an RCG
  technically only holds for \emph{non-combinatory} RCGs, in which the
  arguments of body terms can only contain single variables. Since any
  \emph{combinatory} RCG can be converted into a non-combinatory RCG
  and we only consider non-combinatory RCGs here, this description
  suffices.}

We extend RCGs to PRCGs by annotating each clause $C_k \in P$ with
probabilities $p_k$ such that for all predicates ${A \in N, \,
  \sum_{k:head(C_k)=A} p_k = 1}$. A PRCG defines a distribution over
strings $x$ by sampling from derivations of $S(x)$ according to the
product of probabilities of clauses used in that derivation. This is a
well defined distribution as long as no probability mass is placed on
derivations of infinite length; in this paper, we only consider PRCGs
with derivations of finite length.

\subsection{Learning Model}

\[ \vec{w}_{A_k} \sim Dir(\vec{\alpha}_{A_k})\hspace{3em}
   x_j \underset{iid}{\sim} p_{\text{\textsc{prcg}}}(S(x_j)\,|\,\{w_{A_k}\}) \]
\[  p(\{\vec{w}_{A_k}\}\,|\, \vec{x}, \{\vec{\alpha}_{A_k}\}) \propto
  \prod_j p_{\text{\textsc{prcg}}}(x_j|\{\vec{w}_{A_k}\}) \prod_{A_k}
  p_{\text{\textsc{dir}}}(\vec{w}_{A_k}\,|\,\vec{\alpha}_{A_k}) \]

  \vspace{-1em}
Given a collection of predicates, $\{A_k\}_{k=1}^{K}$, and a
distribution over clauses, $\{\vec{w}_{A_k}\}$, the learning task is
to model a set of outputs, $\{x_j\}_{j=1}^{J}$, as being generated
according to the above distribution. The weights of clauses
with head predicate $A_k$ are drawn from a Dirichlet distribution
defined by $\vec{\alpha}_{A_k}$. Each datum $x_j$ is then drawn
from the resulting PRCG. The probability of a set of weights given our data and Dirichlet prior is determined with Bayes' Rule.

\section{Inference \label{sec:implementation}}

LPNs can be encoded as a restricted subclass of PRISM programs; this
is very similar to how PCFGs are encoded in
PRISM~\cite{DBLP:conf/cl/2000}. Bayesian inference over stochastic
grammars and stochastic logic programs has been an active area of
research in recent decades~\cite{DBLP:journals/etai/Muggleton00,
  cussens2001parameter, DBLP:conf/emnlp/LiangPJK07,
  goldwater2006contextual, johnson2006adaptor}.  Variational inference
is a popular approach in this domain and the one we adopt here by
translating LPNs into PRISM programs and using its built-in
Variational Bayes Expectation-Maximization
algorithm~\cite{sato2008variational}. We modify it for use in an
online setting, using the adaptive learning rate of
\cite{ranganath2013adaptive}.

\section{Experiment}

To evaluate LPNs as a probabilistic model of concept acquisition, we
trained an LPN with the architecture in Figure \ref{fig:architecture}
and $K = 4$ latent predicates on a set of sentences expressing
successor and predecessor relations in numbers between one and
ninety-nine. The training set was the collection of sentences $$X =
\{\text{after}\, \langle n \rangle \, \text{comes} \, \langle n+1
\rangle \,|\, n \in 1,\dots,98\} \cup \{\text{before} \,\langle n
\rangle \, \text{comes} \, \langle n-1 \rangle \,|\, n \in
2,\dots,99\},$$ where $\langle n \rangle$ is the number word
corresponding to $n$. The lexicon was the set of number word
corresponding to $1$ through $19$, the decades $20, \dots, 90$, the
empty string, and the words ``before'', ``after'', and ``comes''. See
Figure~\ref{fig:parseexample}, for a possible LPN derivation of an
example sentence. To approximate the evidence children receive (words
for small numbers naturally occuring more often than for larger ones,
but children also rehearse the entire number sequence), we drew these
sample sentences from a sum of a geometric distribution with parameter
$0.5$ and a uniform distribution. These components were weighted
$75\%$ and $25\%$, respectively. We drew $2000$ examples from this
distribution, holding out a subset of sentences, including those in
Table~\ref{tab:results} for evaluation.

For inference, we used default $\frac{1}{D}$ pseudocounts (where $D$ is the
dimensionality of the Dirichlet distributions). We found that different random
initialization for this experiment did not lead to qualitatively different
results, though further investigation will be necessary to see how
robust the algorithm is to local maxima when fitting LPNs.

We evaluated the learned model by asking for Viterbi ({\it i.e.} maximum {\it a
posteriori}) completions of the last words of each held out test
sentence. Table~\ref{tab:results} shows these completions. The grammar
correctly learns much of the structure of these sentences, including
the difference between sentences starting with ``before'' and
``after'' and the edge cases that relate decade words like ``twenty''
to non-decade words like ``twenty one.''

%% Feel free to include if there's space, but this table may not be necessary.
%%
%%   \begin{table}
%%     \centering
%%     \begin{tabular}{>{\tiny} l >{\tiny} l >{\tiny} l >{\tiny} l }
%%       \multicolumn{2}{>{\tiny}l}{$S$(X Y) $\leftarrow$ $A_1$(X, Y) : 1.0000} &      $A_2$(before, comes) : 0.7316 & $A_4$(after, comes) : 0.9990  \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(X, U), $A_3$(V, Y) : 0.5002} & $A_3$(one, two) : 0.3993 & $A_3$(two, three) : 0.2063 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_3$(Y, V), $A_4$(X, U) : 0.3428} & $A_3$(three, four) : 0.1093 & $A_3$(four, five) : 0.0734 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_4$(X, U) : 0.0796} & $A_3$(five, six) : 0.0502 & $A_3$(six, seven) : 0.0355 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(Y, V), $A_2$(X, U) : 0.0712} & $A_3$(eight, nine) : 0.0290 & $A_3$(seven, eight) : 0.0271 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(Y, X), $A_3$(V, U) : 0.0021} & $A_2$(fifty, fifty) : 0.0375 & $A_2$(thirty, thirty) : 0.0361 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(V, Y), $A_2$(X, U) : 0.0013} & $A_2$(eighty, eighty) : 0.0339 & $A_3$(null, one) : 0.0231 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_2$(V, X), $A_3$(Y, U) : 0.0008} & $A_2$(forty, forty) : 0.0332 & $A_2$(twenty, twenty) : 0.0310 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(V, Y) : 0.0008} & $A_2$(seventy, seventy) : 0.0296 & $A_2$(sixty, sixty) : 0.0274 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_2$(Y, V) : 0.0008} & $A_2$(ninety, ninety) : 0.0260 & $A_3$(eighteen, nineteen) : 0.0064 \\
%%       \multicolumn{2}{>{\tiny}l}{$A_1$(X Y, U V) $\leftarrow$ $A_1$(X, U), $A_4$(Y, V) : 0.0004} & $A_3$(sixteen, seventeen) : 0.0049 & $A_3$(eleven, twelve) : 0.0044 \\
%%       $A_3$(nine, ten) : 0.0044 & $A_3$(thirteen, fourteen) : 0.0044 & $A_3$(fourteen, fifteen) : 0.0039 & $A_3$(ten, eleven) : 0.0034 \\
%%       $A_2$(null, fifty) : 0.0043 & $A_3$(eighty, null) : 0.0030 & $A_3$(seventeen, eighteen) : 0.0030 & $A_3$(nine, sixty) : 0.0025 \\
%%       $A_2$(nine, seventy) : 0.0029 & $A_3$(nine, forty) : 0.0020 & $A_2$(null, thirty) : 0.0022 & $A_2$(nine, ninety) : 0.0022 \\
%%       $A_3$(twelve, thirteen) : 0.0015 & $A_3$(fifteen, sixteen) : 0.0015 & $A_3$(null, comes) : 0.0010 & $A_2$(twenty, after) : 0.0014 \\
%%       $A_2$(sixty, before) : 0.0007 & $A_3$(nineteen, comes) : 0.0005 & $A_4$(nine, thirty) : 0.0010 & \\
%%     \end{tabular}
%%     \caption{The 4 predicate LPN trained to model sentences in the number domain. Rules with insignificant weights are removed. This LPN generates the completions in Table \ref{tab:results}.}
%%     \label{tab:grammar}
%%   \end{table}

\begin{figure}
    \begin{subfigure}[b]{0.45\linewidth}
      \begin{tabular}{>{\footnotesize} l >{\footnotesize} l}
        Question & $K=4$ \\ \hline
        after twenty comes \ldots & twenty one \checkmark \\
        after forty five comes \ldots & forty six \checkmark \\
        after forty nine comes \ldots & forty ten $\times$ \\
        after fifty nine comes \ldots & fifty ten $\times$ \\
        after sixty one comes \ldots & sixty two \checkmark \\
        before twenty three comes \ldots & twenty two \checkmark \\
        before thirty comes \ldots & thirty eighty $\times$ \\
        before thirty eight comes \ldots & thirty seven \checkmark \\
        before forty one comes \ldots & forty \checkmark \\
        before fifty three comes \ldots & fifty two \checkmark \\
        before eighty five comes \ldots & eighty four \checkmark \\
        %% after forty seven comes \ldots & forty eight  \checkmark \\
        %% after sixty three comes \ldots & sixty four \checkmark \\
        %% after sixty four comes \ldots & sixty five \checkmark \\
        %% after sixty five comes \ldots & sixty six \checkmark \\
        %% after sixty nine comes \ldots & sixty ten $\times$ \\
        %% after seventy three comes \ldots & seventy four \checkmark \\
        %% after seventy nine comes \ldots & seventy ten $\times$ \\
        %% after ninety five comes \ldots & ninety six \checkmark \\
        %% before sixty eight comes \ldots & sixty seven \checkmark \\
        %% before seventy two comes \ldots & seventy one \checkmark \\
        %% before seventy three comes \ldots & seventy two \checkmark \\
        %% before ninety two comes \ldots & ninety one \checkmark \\
        %% before ninety three comes \ldots & ninety two \checkmark \\
        %% before ninety five comes \ldots &ninety four \checkmark \\
      \end{tabular}
      \caption{}
      \label{tab:results}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\linewidth]{figures/train_number_net_0006_held_out.pdf}
    \caption{}
    \label{fig:heldoutLL}
  \end{subfigure}
  \caption{(\subref{tab:results}) Viterbi completions of held-out sentences for a 4-predicate 1-layer LPN. (\subref{fig:heldoutLL}) expected log likelihood on a set of held-out test data in a 35-predicate 5-layer LPN with.}
\end{figure}

To inspect visually the learned grammar, we thresholded rules
according to the expected number of times they were used in parsing
the training dataset. Table~\ref{tab:grammar} shows all rules with
expected count above $1e-6$. This reduces from $2669$ to $52$ the
number of significant rules. On inspection, predicate $A_2$ forms
``before'' sentences, predicate $A_4$ forms ``after'' sentences,
predicate $A_3$ is successorship recursively defined over the decades
and ones, and predicate $A_2$ is a category for the decade words.

Our LPN does not learn to how to transition between the last word in a decade
and the next decade (e.g. ``seventy nine'' to ``eighty''). Instead, it
makes the intuitively reasonable generalization that ``seventy nine''
should be followed by ``seventy ten.''

\subsubsection*{Acknowledgments?}

\bibliographystyle{amsplain}
\bibliography{nips_workshop}

\end{document}