### 2014-11-10 drafts/Number.tex
Added some untracked files."

\documentclass{article}

\usepackage[top=2in, bottom=1.5in, left=1in, right=1in]{geometry}
\usepackage{amssymb,amsmath}

\usepackage{algorithm}
\usepackage{algpseudocode}

\usepackage{graphicx}

\usepackage{natbib}

\makeatletter
\newcommand{\xRightarrow}[2][]{\ext@arrow 0359\Rightarrowfill@{#1}{#2}}
\makeatother

\newcommand{\derives}{\xRightarrow{*}}

\title{Understanding number: a case study in learning systems of concepts}
\author{Eyal Dechter}

\begin{document}
\maketitle
\date

\section{Abstract}
How do people learn, represent, and reason about systems of inter-related concepts?

\bibliographystyle{bibliographystyle}
\bibliography{main}

\end{document}