### 2014-11-10 drafts/note1.tex

Notes on meeting with Josh T., Josh R. on number concept learning
8/27/2014

- How do children learn number concepts that go beyond counting and
the cardinality principle?

- In particular, how do children represent and learn the set of
complimentary relations and routines that support number knowledge
over time?

- Relations of interest: 
1) Next word in number list: e.g. Next(``three hundred and forty five'', ``three hundred and forty six'')
2) Approximate magnitude: e.g. AM("eighty six", ~80)
3) Exact cardinatlity: e.g. Exact("eight six", 86)

others: 
1) Is one number greater than another: LexGreater(forty five, twenty seven)
2) Number as digits: Digit(forty five, 45)
3) Approximate forms of addition and multiplication.

- How do we account for different stages of number knowledge development: 
Stage 1: (ages 1-4) pre-cp knowledge, numbers 1-10
Stage 2: (ages 4-6) two-digit numbers
Stage 3: (ages >6) adult like extension to larger numbers